---
title: "DSCI_602_M6_Lab_Distributions_in_R_Cody_York"
author: "Cody York"
date: "10/1/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Create a variable X using seq() function

```{r}

x <- seq(1, 10, length=1000)
x

```

## Plot Gamma Probability Desnsity for X

```{r}
library(tidyverse)
library(dslabs)
library(ggplot2)



plot(x,pgamma(x,shape=2,rate=1),type="l",ylim=ylim,
     main="Gamma Probability Function")



```

## Plot the Emperical Cumulative Distribution of the Air Quality.Ozone Dataset

```{r}

ozone <- airquality$Ozone

ozone.ecdf <- ecdf(ozone)

plot(ozone.ecdf, xlab = 'Sample Quantiles of Ozone', ylab = 'O', 
     main = 'Empirical Cumluative Distribution\nOzone Pollution in New York')

```


## Explore Shape and Rate for the Gamma Distribution
    It seems like increasing the shape is increasing the starting y-value of the function, whereas decreasing the rate is elongating the graph. To decide shape and rate you should look at your data and decide what shape and rate would work best for what you are trying to accomplish with the probability.
```{r}

x <- seq(1, 10, length=1000)

# pgamma(x,shape=2,rate=1)
ylim <- c(0.1, 1)

## Increasing/Decreasing the Shape
plot(x,pgamma(x,shape=2,rate=1),type="l",ylim=ylim,
     main="Gamma Probability Function Original")

plot(x,pgamma(x,shape=1,rate=1),type="l",ylim=ylim,
     main="Gamma Probability Function with Shape 1")

plot(x,pgamma(x,shape=3,rate=1),type="l",ylim=ylim,
     main="Gamma Probability Function with Shape 3")

plot(x,pgamma(x,shape=5,rate=1),type="l",ylim=ylim,
     main="Gamma Probability Function with Shape 5")

# Increasing/Decreasing the Rate
plot(x,pgamma(x,shape=2,rate=0.5),type="l",ylim=ylim,
     main="Gamma Probability Function with Rate 0.5")

plot(x,pgamma(x,shape=2,rate=1.5),type="l",ylim=ylim,
     main="Gamma Probability Function with Rate 2")

plot(x,pgamma(x,shape=2,rate=3),type="l",ylim=ylim,
     main="Gamma Probability Function with Rate 3")

# Increasing/Decreasing Both

plot(x,pgamma(x,shape=3,rate=0.5),type="l",ylim=ylim,
     main="Gamma Probability Function with Shape 3 and Rate 0.5")

```
