---
title: "DSCI602_M3Lab_Cody_York"
author: "Cody York"
date: "2022-09-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1 Calculate Expected value and variance for a discrete varaible

```{r}
## Define a Vector
x <- 1:12

## Define the Probability
PDF <- dbinom(x, size=12, prob=.75)
print('The probabilities are:')
print(PDF)
print('The sum of the probabilities is:')
sum(PDF)

## Compute E(X)
expected_value <- sum(x*PDF);
print('The expected value is: ')
print(expected_value)

## Compute Var(X)
variance <- sum((x-expected_value)^2*PDF)
print('The varaince is: ')
print(variance)

```

## 2 Calculate the expected value and variance for a continuous variable

```{r}
## Define a Vector
x <- 1:10

## Define the Equations for a continuous variable
f <- function(x) 3/(x^4)
g <- function(x) x * f(x)
h <- function(x) x^2 * f(x)

# Compute E(x)
expected_value <- integrate(g,
                            lower = 1,
                            upper = Inf
                          )$value
print("The expected value is: ")
print(expected_value)

# Compute Variance
variance <- integrate( h,
                 lower = 1,
                 upper = Inf
                )$value - (expected_value)^2
print("The variance is: ")
print(variance)

```

## 3 Matrix Questions

```{r}
## Create the Matrices

A_values <- c(5,0,0,5)
A <- matrix(A_values, nrow=2, ncol=2)
print("Matrix A: ")
print(A)

B_values <- c(4,1,3,0,2,-1)
B <- matrix(B_values, nrow=3, ncol=2, byrow=TRUE)
print("Matrix B: ")
print(B)

```

## 3a Can you add the above matrixes?
    
    These two arrays cannot be added because they do not match in size. We cannot add a matrix of 2x2 and 3x2 since they have a different row count.

```{r}
## Try to Add
## Commented out code since this causes a failure due to above reason.

# matrix_sum <- A + B
# print("Try to sum the matrices: ")
# print(matrix_sum)
```

## 3b Can you find AB or BA?
    
    A * B cannot be done due to the dimensions not being compatible since A's Columns do not match B's Rows.
    
    B * A can be multiplied since they have a matching column to row count.

```{r}
## Try to multiply
## Multiply using matrix %*% matrix 

# ab_matrix_multiply <- A %*% B
# print("AB Matrix Multiply: ")
# print(ab_matrix_multiply)


ba_matrix_multiply <- B%*%A
print("BA MAtrix Multiply: ")
print(ba_matrix_multiply)

```

## 3c What is the transpose of B?

```{r}

## Find the Transpose of B using t()

b_transpose <- t(B)
print("The transpose of B is: ")
print(b_transpose)
```

## 3d Find the Determinent of BtB

```{r}
## Use the det function in r. In the function multiply the transpose of B by B
det_of_btb <- det(t(B) %*% B)
print("The Determinent is: ")
print(det_of_btb)
```
