---
title: "Cody_York_M11_lab"
author: "Cody York"
date: "2022-11-07"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Imports
    This is where I am keeping all my R library import statements.

```{r}
library(readr)
library(plotrix)

```

## Read the Dataset
    For this step I am using the readr library to read in the csv containing the data and viewing it to make sure it is coming correctly. I will also get the population mean since we will need that in the future.

```{r}

Soil_Organic_Carbon <- read_csv("Soil Organic Carbon.csv")
#View(Soil_Organic_Carbon)

pop_mean <- mean(Soil_Organic_Carbon$`0-5cm`)
```

## Sample 50 rows and for 100 times calulate the 95% Confidence Level
    For the first part of this task I will be using the sampling method we learned form t he last R lab. 
    
  1. I first set the seed so that the randomness won't be different every time for testing the code.
  2. Next, I create a variable to hold the number of samples I need.
  3. Next, I set a variable to the number of runs.
  4. Now, I will create a function that will run 100 times to get the data we need.
  5. Then, I get the samples from the imported data set above using the sample() command in R.
  6. After this I will calculate the z-value using the qnorm() function and the alpha of 95% confidence interval which is .025.
  7. I will then get the mean and standard deviation of the sample.
  8. Next, I will calculate the margin of error. This is the z-value times the standard deviation divided by the square root of the sample size.
  9. After getting those pieces of information, I will calculate the lower and upper bounds of the confidence interval by subtracting and adding the margin of error to the sample mean.
  10. I will then run the function 100 times using the replicate function to get the data.
  11. Next I do a transpose on the data to get the three observations I need.
    
```{r}

set.seed(111)

number_of_samples <- 50

runs <- 100

population_mean <- 

one.trial <- function(){
  Soil_Organic_Carbon_50_Sample <- Soil_Organic_Carbon[
    sample(nrow(Soil_Organic_Carbon),number_of_samples),2
  ]
  positive_z <- qnorm(.975)
  negative_z <- qnorm(.025)
  sample_mean <- mean(Soil_Organic_Carbon_50_Sample$`0-5cm`)
  standard_deviation <- sd(Soil_Organic_Carbon_50_Sample$`0-5cm`)
  margin_of_error <- positive_z * (standard_deviation/sqrt(number_of_samples))
  lower_bound <- sample_mean - margin_of_error
  upper_bound <- sample_mean + margin_of_error
  middle_of_bounds <- (lower_bound + upper_bound) / 2
  confidence_intervals <- c(middle_of_bounds, upper_bound, lower_bound)
  return(list(middle_of_bounds, upper_bound, lower_bound))
}

sample_50_confidence_interval <- replicate(runs, one.trial())

final_sample_50_confidence_interval <- as.data.frame(
  t(sample_50_confidence_interval))


```

## Graph the 50 sample confidence intervals
  Now I will graph the confidence intervals collected above for the 50 samples.
  1. First, I will set the sample index to the number of runs form above.
  2. Next, I will use plotCI() to plot the middle, upper, and lower confidence intervals.
  3. Then I will add the population mean to the observations in the plotCI().

```{r}

sample_index <- c(1:runs)

plotCI(sample_index, 
       as.numeric(final_sample_50_confidence_interval$V1),
       as.numeric(final_sample_50_confidence_interval$V2),
       as.numeric(final_sample_50_confidence_interval$V3),
       lwd = 0.5,
       col = "red",
       scol = "green",
       ylab = c("CI"),
       xlab = c("Sample Index")
       )
plotCI(sample_index, 
       rep(as.integer(pop_mean), runs), 0, 0, col = "blue", add = TRUE)


```
## Sample 15 rows and for 100 times calulate the 95% Confidence Level
    I will now do the same as above but only use 15 rows as a sample instead of 50.
    
```{r}

set.seed(111)

number_of_samples <- 15

runs <- 100

population_mean <- 

one.trial <- function(){
  Soil_Organic_Carbon_15_Sample <- Soil_Organic_Carbon[
    sample(nrow(Soil_Organic_Carbon),number_of_samples),2
  ]
  positive_z <- qnorm(.975)
  negative_z <- qnorm(.025)
  sample_mean <- mean(Soil_Organic_Carbon_15_Sample$`0-5cm`)
  standard_deviation <- sd(Soil_Organic_Carbon_15_Sample$`0-5cm`)
  margin_of_error <- positive_z * (standard_deviation/sqrt(number_of_samples))
  lower_bound <- sample_mean - margin_of_error
  upper_bound <- sample_mean + margin_of_error
  middle_of_bounds <- (lower_bound + upper_bound) / 2
  confidence_intervals <- c(middle_of_bounds, upper_bound, lower_bound)
  return(list(middle_of_bounds, upper_bound, lower_bound))
}

sample_15_confidence_interval <- replicate(runs, one.trial())

final_sample_15_confidence_interval <- as.data.frame(
  t(sample_15_confidence_interval))

```

## Graph the 15 sample confidence intervals
    I will now graph the 15 samples like I did for the above 50 samples.
    
```{r}

sample_index <- c(1:runs)

plotCI(sample_index, 
       as.numeric(final_sample_15_confidence_interval$V1),
       as.numeric(final_sample_15_confidence_interval$V2),
       as.numeric(final_sample_15_confidence_interval$V3),
       lwd = 0.5,
       col = "red",
       scol = "green",
       ylab = c("CI"),
       xlab = c("Sample Index")
       )
plotCI(sample_index, 
       rep(as.integer(pop_mean), runs), 0, 0, col = "blue", add = TRUE)



```