---
title: "DSCI602_M2Lab_Cody_York"
author: "Cody York"
date: "2022-09-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Imports

```{r}
library(matrixStats)
library("DescTools")
```

## 2 3X5 Matrix

```{r}
my.matrix <- matrix(1:15, nrow = 3, ncol = 5)
print(my.matrix)
```

## 2a row means and column means

```{r}
my.matrix.rowMeans <- rowMeans(my.matrix)
print(my.matrix.rowMeans)

my.matrix.colMeans <- colMeans(my.matrix)
print(my.matrix.colMeans)
```

## 2b row standard deviations and column standard deviations

```{r}
my.matrix.rowSds <- rowSds(my.matrix)
print(my.matrix.rowSds)

my.matrix.colSds <- colSds(my.matrix)
print(my.matrix.colSds)
```

## 3 Load any R dataset and calculate mean,median,mode and maximum

```{r}
iris_data <- iris
names(iris_data)
class(iris_data$Petal.Length)
petal.length <- iris_data$Petal.Length

petal.length.mean <- mean(petal.length)
print(petal.length.mean)

petal.length.median <- median(petal.length)
print(petal.length.median)

petal.length.mode <- Mode(petal.length)
print(petal.length.mode)

petal.length.max <- max(petal.length)
print(petal.length.max)
```

# 4 Feedback from Textbook

   I feel that after a few moments of looking through the code and reading what was being said about the code block I could come to understand. I think the difficulty of the code is suitable to someone who knows programming languages but maybe not R. I like that the textbook explains the methods being used and why. For the syntax of R I can always review the code blocks in the textbook. I think overall this chapter in the textbook was medium difficulty.