---
title: "DSCI_602_Project_2_Cody_York"
author: "Cody York"
date: "2022-12-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyverse)
```

## Data Ingestion and Pre-processing
    
    The below R code will ingest the data and do any cleanup or preprocessing required. The datasets that will be used for statistics and graphing will also be created here.
    
```{r}

COVID_19_Deaths <- read_csv("Provisional_COVID-19_Deaths_by_Sex_and_Age.csv")

Covid_19_Total_Deaths <- COVID_19_Deaths[1,10]

Total_Deaths <- COVID_19_Deaths[1,11]

COVID_19_Deaths_by_Age <- COVID_19_Deaths %>%
                            filter(COVID_19_Deaths$State == "United States" 
                                   & COVID_19_Deaths$`Age Group` != "All Ages"
                                   & COVID_19_Deaths$Sex == "All Sexes"
                                   & COVID_19_Deaths$Group == "By Total"
                                   ) %>%
                            select(`Age Group`, `COVID-19 Deaths`,
                                   `Total Deaths`)

COVID_19_Deaths_by_Age_DF <- data.frame(COVID_19_Deaths_by_Age)

COVID_19_Deaths_by_Sex <- COVID_19_Deaths %>%
                          filter(COVID_19_Deaths$State == "United States" 
                                   & COVID_19_Deaths$`Age Group` == "All Ages"
                                   & COVID_19_Deaths$Sex != "All Sexes"
                                   & COVID_19_Deaths$Group == "By Total"
                                   ) %>%
                            select(Sex, `COVID-19 Deaths`,`Total Deaths`)


COVID_19_Deaths_by_Sex_DF <- data.frame(COVID_19_Deaths_by_Sex)

```

## Graphing if Age Increases Chance of Death by Covid-19 
    
    The below code will graph the different age groups into a barplot with the calculated average Covid-19 Deaths. I am multiplying by 100 so we can see a percentage in the graph. This average was calculated by taking each age category and dividing it by the earlier created variable Covid_19_Total_Deaths. I have also created the average total deaths by age group using the the earlier created variable Total_Deaths. I have also created an age_groups variable so I can get only the age_groups I want to show in the graph. I have also calculated the mean and standard deviation of the data to be used later.

```{r}
  
  Covid_19_Avg_Deaths_By_Age <- COVID_19_Deaths_by_Age_DF %>%
                                  group_by(Age.Group) %>%
                                  filter(COVID.19.Deaths != "NA") %>%
                                  summarise(
                                    deaths_by_age = sum(COVID.19.Deaths),
                                    total_deaths_by_age = sum(Total.Deaths),
                                    avg_deaths_by_age = round 
                                    (
                                      deaths_by_age / Covid_19_Total_Deaths,
                                      digits = 4
                                    ),
                                    avg_deaths_by_age_from_total = round
                                    (
                                      deaths_by_age / Total_Deaths,
                                      digits = 4
                                    )
                                  )
  
  age_groups <- c("0-17 years", "18-29 years", "30-39 years", "40-49 years",
                  "50-64 years", "65-74 years", "75-84 years", 
                  "85 years and over")
  
  Covid_19_Avg_Deaths_By_Age_DF <- data.frame(Covid_19_Avg_Deaths_By_Age) %>%
                                   filter(Age.Group %in% age_groups)
  
  std_dev_of_Covid_19_Avg_Deaths_By_Age <- sd(
    as.numeric(
      unlist(
        Covid_19_Avg_Deaths_By_Age_DF$avg_deaths_by_age
        )
      )
    )
  
  mean_of_Covid_19_Avg_Deaths_By_Age <- mean(
    as.numeric(
      unlist(
        Covid_19_Avg_Deaths_By_Age_DF$avg_deaths_by_age
        )
      )
    )

 barplot(as.numeric(
   unlist(Covid_19_Avg_Deaths_By_Age_DF$avg_deaths_by_age))*100,
          ylim = c(0,100),
          names.arg = Covid_19_Avg_Deaths_By_Age_DF$Age.Group,
          main = "Covid-19 Deaths Average % by Age Group",
          cex.names = 0.5,
          xlab = "Age Group",
          ylab = "Average Deaths by %",
          border = "black",
          col = rainbow(6)
         ) 
```

## Graphing Covid-19 Age Deaths Based on Total Deaths 
    The below code will graph the different age groups into a barplot with the calculated average Total Deaths by age. I am multiplying by 100 so we can see a percentage in the graph.

```{r}

 barplot(as.numeric(
   unlist(Covid_19_Avg_Deaths_By_Age_DF$avg_deaths_by_age_from_total))*100,
          ylim = c(0,100),
          names.arg = Covid_19_Avg_Deaths_By_Age_DF$Age.Group,
          main = "Covid-19 Death Average % by Age Group from Total Deaths",
          cex.names = 0.5,
          xlab = "Age Group",
          ylab = "Average Deaths by %",
          border = "black",
          col = rainbow(6)
         ) 


```

## Graphing Total Deaths by Age Group

  The below code will graph the different age groups into a barplot with the Total Deaths by age. This graph is simply for informational purposes.

```{r}

 barplot(as.numeric(
   unlist(Covid_19_Avg_Deaths_By_Age_DF$total_deaths_by_age)),
          names.arg = Covid_19_Avg_Deaths_By_Age_DF$Age.Group,
          main = "Total Deaths by Age Group",
          cex.names = 0.5,
          xlab = "Age Group",
          ylab = "Total Deaths",
          border = "black",
          col = rainbow(6)
         ) 

```

## Graphing if Gender Increases Chance of Death by Covid-19 
  The below code will graph the different genders into a barplot with the calculated average Covid-19 Deaths. I am multiplying by 100 so we can see a percentage in the graph. This average was calculated by taking each gender category and dividing it by the earlier created variable Covid_19_Total_Deaths. I have also created the average total deaths by age group using the the earlier created variable Total_Deaths. I have also calculated the mean and standard deviation of the data to be used later.

```{r}

  Covid_19_Avg_Deaths_By_Gender <- COVID_19_Deaths_by_Sex_DF %>%
                                  group_by(Sex) %>%
                                  filter(COVID.19.Deaths != "NA") %>%
                                  summarise(
                                    deaths_by_sex = sum(COVID.19.Deaths),
                                    avg_deaths_by_sex = round 
                                    (
                                      deaths_by_sex / Covid_19_Total_Deaths,
                                      digits = 4
                                    ),                                                                             avg_deaths_by_sex_from_total = round
                                    (
                                      deaths_by_sex / Total_Deaths,
                                      digits = 4
                                    ),
                                    total_deaths_by_sex = sum(Total.Deaths)
                                  )

 Covid_19_Avg_Deaths_By_Gender_DF = data.frame(Covid_19_Avg_Deaths_By_Gender)
 
   std_dev_of_Covid_19_Avg_Deaths_By_Gender <- sd(
    as.numeric(
      unlist(
        Covid_19_Avg_Deaths_By_Gender_DF$avg_deaths_by_sex
        )
      )
    )
   
  mean_of_Covid_19_Avg_Deaths_By_Gender <- mean(
    as.numeric(
      unlist(
        Covid_19_Avg_Deaths_By_Gender_DF$avg_deaths_by_sex
        )
      )
    )
 
  barplot(as.numeric(
   unlist(Covid_19_Avg_Deaths_By_Gender_DF$avg_deaths_by_sex))*100,
          ylim = c(0,100),
          names.arg = c("Male", "Female"),
          main = "Covid-19 Deaths Average % by Sex",
          cex.names = 1,
          xlab = "Gender",
          ylab = "Average Deaths by %",
          border = "black",
          col = c("blue", "pink")
         ) 
 
```

## Graphing Covid-19 Gender Deaths Based on Total Deaths
  The below code will graph the different gender groups into a barplot with the calculated average Total Deaths by gender. I am multiplying by 100 so we can see a percentage in the graph.
```{r}

  barplot(as.numeric(
   unlist(Covid_19_Avg_Deaths_By_Gender_DF$avg_deaths_by_sex_from_total))*100,
          ylim = c(0,100),
          names.arg = c("Male", "Female"),
          main = "Covid-19 Death Average % by Gender from Total Deaths",
          cex.names = 1,
          xlab = "Gender",
          ylab = "Average Deaths by %",
          border = "black",
          col = c("blue", "pink")
         ) 

```

## Graphing Total Deaths by Gender
  The below code will graph the different genders into a barplot with the Total Deaths by gender. This graph is simply for informational purposes.

```{r}

  barplot(as.numeric(
   unlist(Covid_19_Avg_Deaths_By_Gender_DF$total_deaths_by_sex)),
          names.arg = c("Male", "Female"),
          main = "Total Deaths by Gender",
          cex.names = 1,
          xlab = "Gender",
          ylab = "Total Deaths",
          border = "black",
          col = c("blue", "pink")
         ) 

```

## Hypothesis Testing
    This section contains my hypothesis testing in R for the previous statistics. I have included a t_test for both age average and gender average. I have also created an F test for both datasets and a chi-squared test for each dataset. I have also included a one way and two way anova. 
```{r}

t.test(Covid_19_Avg_Deaths_By_Age_DF$avg_deaths_by_age, 
       mu = mean_of_Covid_19_Avg_Deaths_By_Age, 
       alternative = "less")

t.test(Covid_19_Avg_Deaths_By_Gender_DF$avg_deaths_by_sex, 
       mu = mean_of_Covid_19_Avg_Deaths_By_Gender, 
       alternative = "less")

var.test(as.numeric(unlist(Covid_19_Avg_Deaths_By_Age_DF$avg_deaths_by_age)),
         as.numeric(unlist(Covid_19_Avg_Deaths_By_Gender_DF$avg_deaths_by_sex)))

chisq.test(Covid_19_Avg_Deaths_By_Age_DF$avg_deaths_by_age)

chisq.test(Covid_19_Avg_Deaths_By_Gender_DF$avg_deaths_by_sex)

```